---
title: "Global Crop Yield"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dpi = 180)
library(tidyverse)
theme_set(theme_light())
```

### Inspired by Julia Silge
For this week's tidytuesday data on [global crop yields around the world] (https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-09-01/readme.md), we want to create a model that will estimate how crop yields are changing around the world. There are 5 different data, but we will not look at all of them. 

## Data Exploration
We first get the data from the link above. For this analysis, we'll only use key_crop_yields and land_use data.
```{r, include=FALSE}
key_crop_yields <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-01/key_crop_yields.csv')
land_use <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-01/land_use_vs_yield_change_in_cereal_production.csv')

```

Again, we will only be looking on some countries which will be done based on  population.

```{r, include=FALSE}
top_countries <- 
  land_use %>% 
  janitor::clean_names() %>%
  filter(!is.na(code),
         entity != 'World') %>% 
  group_by(entity) %>% 
  filter(year == max(year)) %>% 
  ungroup() %>% 
  slice_max(total_population_gapminder, n = 30) %>% 
  pull(entity)
```

We'll then combine this table with key_crop_yields and will only look at some crops that we're interested in.

```{r}
tidy_yields <- 
  key_crop_yields %>%
  janitor::clean_names() %>%
  pivot_longer(wheat_tonnes_per_hectare:bananas_tonnes_per_hectare,
    names_to = "crop", values_to = "yield") %>%
  mutate(crop = str_remove(crop, "_tonnes_per_hectare")) %>%
  filter(crop %in% c("wheat", "rice", "maize", "barley"),
    entity %in% top_countries,
    !is.na(yield))

```

```{r}
tidy_yields %>% 
  ggplot(aes(x = year, y = yield, color = crop)) + 
  geom_line(alpha = 0.7, size = 1.5) + 
  geom_point() + 
  facet_wrap(~entity) + 
  labs(x = NULL, y = "yield (tonnes/hectare)")
```

## Create Models
```{r}
library(tidymodels)

tidy_lm <- 
  tidy_yields %>% 
  nest(yields = c(year, yield)) %>% 
  mutate(model = map(yields, ~ lm(yield ~ year, data = .x)))

slopes <- tidy_lm %>% 
  mutate(coefs = map(model, tidy)) %>% 
  unnest(coefs) %>% 
  filter(term == "year") %>% 
  mutate(p.value = p.adjust(p.value))
```

## Explore Results
```{r}
library(ggrepel)
slopes %>% 
  ggplot(aes(x = estimate, y = p.value, label = entity)) +
  geom_vline(xintercept = 0, lty = 2, size = 1.5, alpha = 0.7,
             color = 'gray') + 
  geom_point(aes(color = crop), alpha = 0.8, size = 2.5,
             show.legend = FALSE) + 
  geom_text_repel(size = 3, family = "IBMPlexSans") +
  facet_wrap(~ crop) +
  scale_y_log10()
```

