---
title: "food-consumption"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Inspired by Julia Silge
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dpi = 180,
                      fig.width = 12, fig.height = 5)

library(tidyverse)
theme_set(theme_light())
```

This time round, we want to predict which countries are Asian countries, based on their patterns of food consumption in the eleven categories from [food consumption and CO2 emissions.](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-18/readme.md)


```{r}
food_consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv')
food_consumption
```
We have a long, tidy data which needs to be made wide for modeling. To identify the continent which each country resides in, we can use countrycode package. We will also create a new variable 'asia' for prediction that tells us whether a country is in Asia or not.

```{r}
library(janitor)
library(countrycode)
food <- food_consumption %>% 
  select(-co2_emmission) %>% 
  pivot_wider(names_from = food_category,
              values_from = consumption) %>% 
  clean_names() %>% 
  mutate(continent = countrycode(country,
                                 origin = "country.name",
                                 destination = "continent")) %>% 
  mutate(asia = case_when(continent == "Asia" ~ "Asia",
                          TRUE ~ "Other")) %>% 
  select(-country, -continent) %>% 
  mutate_if(is.character, factor)
food
```
We are left with quite a small dataset, but we will try to model this data to see how we can do hyperparameters tuning.

```{r}
library(GGally)
ggscatmat(food, columns = 1:11, color = "asia", alpha = 0.7)
```
Notice how important rice is! Also see how the relationships between different food categories is different for Asian and non-Asian countries; a tree-based model like a random forest is good as learning interactions like this.

## Hyperparameters tuning
- Create a set of bootstrap resamples to use for tuning
- Create a model specification for a random forest where we will tune mtry (# of predictors to sample at each split) and min_n (# of observations needed to keep splitting nodes). 

```{r}
library(tidymodels)
set.seed(1234)
food_boot <- bootstraps(food, times = 30)
food_boot
```

```{r}
rf_spec <- rand_forest(
  mode = "classification",
  mtry = tune(),
  trees = 1000,
  min_n = tune()) %>% 
  set_engine("ranger")

rf_spec
```

Here, we will use grid search to train many models. Furthermore, we can do parallel processing to speed up our training.
```{r}
doParallel::registerDoParallel()

rf_grid <- tune_grid(
  asia ~. ,
  model = rf_spec,
  resamples = food_boot)

rf_grid

```

We can check our results
```{r}
rf_grid %>% 
  collect_metrics()
```

And to see which model performed best in terms of 'accuracy' or 'roc_auc'.
```{r}
rf_grid %>% 
  show_best("roc_auc")
```


```{r}
world <- map_data("world")

food_country <- food_consumption %>% 
  left_join(world %>% mutate(country = region), by = 'country')
```

```{r}
food_country %>% 
  filter(food_category == "Pork") %>% 
  ggplot(aes(long, lat, group = group))+
  geom_polygon(aes(fill = consumption), color = "white")+
  scale_fill_viridis_c(option = "C")
```

```{r, echo = FALSE}
shinyAppFile("/Users/indrikwijaya/Documents/tidytuesday/2020-02-18/app.R"
)
```
